---
title: "Costumers Data Analysis"
author: "Karim Mezhoud"
date: "12/17/2018"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
# Load dataset from url
basket <- read.csv("http://www.evodevo.it/wp-content/uploads/2018/12/basket.csv")
# Backup the dataset
#write.csv(basket, file = "basket_bkp.csv")
basket
```

## Transactions list
```{r}
#library(plyr)

transaction_list_costumer <- plyr::ddply(basket,c("Member_number","Date"),
                                function(df1)paste(df1$itemDescription,
                                                   collapse = ","))
colnames(transaction_list_costumer) <- c("Member_number","Date","Baskets")
transaction_list_costumer
```

## How many members? and How many items? 
```{r}
library(dplyr)
# How many member
base::table(basket$Member_number) %>% length()
# 3898

# How many items
base::table(basket$itemDescription) %>% length()
# 167
```


## Ranking the most important Customers by the number of transactions
```{r}
library(dplyr)
# How many transactions by Member, the data will be arranged by descending
basket %>% group_by(Member_number) %>% 
  arrange(Member_number) %>% 
  dplyr::summarise(N_operation=n()) %>% 
  arrange(desc(N_operation))
```

## The 10 most important customers are:
```{r}
Costumers_10 <- basket %>% group_by(Member_number) %>% 
  arrange(Member_number) %>% 
  dplyr::summarise(N_operation=n()) %>% 
  arrange(desc(N_operation)) %>%
  head(10) %>%
  tibble::as.tibble()
Costumers_10
```

## Extract the transactions list of the 10 most important costumers

```{r}

target <- c("3180", "2015", "3050", "3737", "2271", "2433", "2625", "3915", "3872", "2394")
transaction_list_costumer_10 <- transaction_list_costumer %>%
  filter( Member_number %in% target)
transaction_list_costumer_10
```

# get the baskets of the 10 most important costumers
```{r}
#set column Member_number of dataframe transaction_list 
transaction_list_costumer_10$Member_number <- NULL
#set column Date of dataframe transaction_list
transaction_list_costumer_10$Date <- NULL
#Rename column to Baskets
colnames(transaction_list_costumer_10) <- c("Baskets")
#Show Dataframe transactionData
transaction_list_costumer_10
```


```{r}
library(arules)
# Write the Baskets list to a file
write.csv(transaction_list_costumer_10,"transactions_list_costumer_10.csv", quote = FALSE, row.names = TRUE)
# load the Baskets list as a basket format using arules package
tr_list_costumer_10 <- read.transactions('transactions_list_costumer_10.csv', format = 'basket', sep=',')

tr_list_costumer_10
```

## Summary of the Baskets of the 10 most imortant costumers
```{r}
summary(tr_list_costumer_10)
```
The summary shows that the baskets with **3** items are the most frequent: **50** transactions.

## The most 10 important items for the most 10 important costumers

```{r}
library(RColorBrewer)
#par(mfrow=c(2,1))
arules::itemFrequencyPlot(tr_list_costumer_10,topN=10,type="absolute",col=brewer.pal(8,'Pastel2'), main="Absolute Product Frequency Plot for the 10 most important Costumers")
arules::itemFrequencyPlot(tr_list_costumer_10,topN=10,type="relative",col=brewer.pal(8,'Pastel2'), main="Relative Product Frequency Plot for the 10 most important Costumers")
```

The `whole milk` and `other vegetables`  remain at the first rank. The `yogurt` transactions progress from **5th** position to the **3rd**.  Also we observe a progress for `Sausage` (9->6th), but a considerable regress of the `root vegetables` from **6th** to **10th** position.

## Apply ```Arules``` algothimes to predict Rules that associate items

```{r}
#Min Support as 0.0001, confidence as 0.9 and maximum of 3 (where ther the most number of transaction (itemset/transaction)) products.
association.rules_Costumer_10 <- arules::apriori(tr_list_costumer_10, parameter = list(supp=0.0001, conf=0.9, maxlen=10))
```

With 0.0001 support, 0.9 of confidence and 3 maximums items per basket, we obtain **17690** rules.

## Summary of the Rules of the 10 most important items specific to the 10 most important costumers

```{r}
# summary rules of min Support as 0.0001, confidence as 0.8 and maximum of 10 products.
summary(association.rules_Costumer_10)
```

## Inspect the association of the rules

```{r}
inspect(association.rules_Costumer_10[1:15])
```
+ The inspection shows that the most 10 important costumers got transactions with small basket mainly with 3 items. 

+ There is no a complexe association between the 10 most important items.

+ The costumers bougth directly, mainly 3 items whitout running through supermarket department.


```{r}
library(arulesViz)
# Filter rules with confidence greater than 0.4 or 40%
subRules_costumer_10<-association.rules_Costumer_10[quality(association.rules_Costumer_10)$confidence>0.4]
#Plot SubRules
plot(subRules_costumer_10)
plot(subRules_costumer_10,method="two-key plot")
```

+ This plot shows that the most frequent fraction of transactions that contain  3 or 4 items (purple, blue) (support).

## Interactive visualisation of the Rules of the most 10 important items for the best 1O costumers

```{r}
#top10subRules_costumer_10 <- head(subRules_costumer_10, n = 16, by = "confidence")
plot(subRules_costumer_10[1:20], method = "graph",  engine = "htmlwidget")
```

**Interpretation**

+ The inspection shows that the most 10 important costumers got transactions with small basket mainly with 3 items. 

+ There is no a complexe association between the 10 most important items.

+ The costumers bougth directly, mainly 3 items whitout running through supermarket department.

# Individual Rule Representation

```{r}
# Filter top 10 rules with highest lift
subRules2_costumer_10<-head(subRules_costumer_10, n=10, by="lift")
plot(subRules2_costumer_10, method="paracoord") #, control = list(reorder = TRUE))
```

** Interpretation$**

+ There is not particular significant rule for the 10 most important costumers.

+ The 10 most important costumers buy small basket (3 or 4 items) without running through supermarket department
